MPA Gap Analysis: IUCN and AquaMaps

### 2016-03-01 mpa_analysis/mpa_gap_analysis.Rmd
let's get some of this up on line...

``` {r set_up_base_raster}
### DO WE NEED THIS? Use area at 0.5 degree level; most cells this will be pretty close.
### base raster will contain 0.01 degree cells from -180 to +180,
### and -90 to +90, with the associated area of each cell.
### Save it as a .tif in GitHub - about 50 
# rast_base_file <- file.path(dir_git, 'mpa_analysis/rast_base.tif')
# 
# if(!file.exists(rast_base_file)) {
#   ### import AquaMaps half-degree loiczid raster - for extents and such
#   rast_loiczid_file   <- file.path(dir_git, 'data/rasters/loiczid_raster.grd')
#   rast_loiczid        <- raster(rast_loiczid_file)
#   names(rast_loiczid) <- 'loiczid'
# 
#   rast_empty <- raster(ext = extent(rast_loiczid), resolution = 0.01)
#   rast_base  <- raster::area(rast_empty) ### takes a while...
#   
#   writeRaster(rast_base, rast_base_file,
#               progress = 'text', overwrite = TRUE)
# } else {
#   rast_base <- raster(rast_base_file)
# }

@@

dir_rast <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters')
rast_mpa_mar_file <- file.path(dir_rast, 'rast_mpa_mar.tif')
rast_mar_file <- file.path(dir_rast, 'rast_marine.tif')

if(!file.exists(rast_mpa_mar_file)) {
  rast_mpa_file <- file.path(dir_rast, 'WDPA_poly_Jan2015_mpa1.tif')

@@

rast_loiczid_01_file <- file.path(dir_rast, 'rast_loiczid_01.tif')
if(!file.exists(rast_loiczid_01_file)) {
  rast_loiczid_01 <- resample(rast_loiczid, rast_mpa_mar, 
                            method = 'ngb',
                            progress = 'text',
                            filename = rast_loiczid_01_file)
# writeRaster(rast_loiczid_01, rast_loiczid_01_file, progress = 'text')
} else {
  rast_loiczid_01 <- raster(rast_loiczid_01_file)
}
```

@@

### Count marine cells in each LOICZID cell
mar_df_file <- file.path(dir_git, 'mpa_analysis/loiczid_marine.csv')
if(!file.exists(mar_df_file)) {
  is_marine_df <- zonal(rast_mar, 
                        z = rast_loiczid_01, 
                        fun = 'count',
                        progress = 'text') %>%
    as.data.frame() %>%
    rename(loiczid = zone)
  write_csv(is_marine_df, mar_df_file)
}

### Count all protected cells in each LOICZID cell
### NOTE: loses status_yr values - so filter by year for each iteration
yr_list <- c('y1990', 'y1995', 'y2000', 'y2005', 'y2010', 'y2014', 'y2015')
mpa_csvs_missing <- !file.exists(file.path(dir_git, sprintf('mpa_analysis/loiczid_mpa_%s.csv', yr_list)))

if(any(mpa_csvs_missing)) {
  yr_list1 <- yr_list[mpa_csvs_missing]
  message('Creating marine protected area rasters for years:')
  message('  ', paste(yr_list, collapse = ', '))
  for (yr in yr_list1) { # yr <- yr_list1[1]
    message('filtering MPAs to those designated prior to ', str_replace(yr, 'y', ''))
    rast_is_mpa <- rast_mpa_mar
    data_year <- as.integer(str_replace(yr, 'y', ''))
    values(rast_is_mpa)[values(rast_mpa_mar) >= data_year] <- NA
      ### NOTE: 2015 MPA dataset includes up to 2014; 2014 dataset goes
      ###   through 2013; etc.
    message('counting protected cells per LOICZID for ', yr)
    is_mpa_df <- zonal(rast_is_mpa, 
                       z = rast_loiczid_01, 
                       fun = 'count',
                       progress = 'text') %>%
      as.data.frame() %>%
      rename(loiczid = zone)
    write_csv(is_mpa_df, 
              file.path(dir_git, sprintf('mpa_analysis/loiczid_mpa_%s.csv', yr)))
  }
}
```

@@

dir_am_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/v2015/csv/')

hcaf <- read_csv(file.path(dir_am_data, 'hcaf_truncated.csv')) %>%
  select(loiczid, cellarea, oceanarea)

cell_df <- read_csv(file.path(dir_git, 'mpa_analysis/loiczid_marine.csv'),
  col_types = 'dd') %>%
  rename(marine_ct = count) %>%
  left_join(hcaf, by = 'loiczid') %>%
  mutate(marine_pct  = marine_ct/2500,
         marine_area = marine_pct * cellarea) %>%
  select(-marine_ct)

for (yr in yr_list) { # yr <- yr_list[1]
  fn <- file.path(dir_git, sprintf('mpa_analysis/loiczid_mpa_%s.csv', yr))
  df <- read_csv(fn, col_types = 'dd')
  names(df)[2] <- yr
  cell_df <- cell_df %>%
    left_join(df, by = 'loiczid')
  cell_df[ , paste(yr, '_pct', sep = '')]  <- cell_df[ , yr]/2500
  cell_df[ , paste(yr, '_area', sep = '')] <- cell_df[ , paste(yr, '_pct', sep = '')] * cell_df$cellarea

@@

@@

### group by speciesid, then summarize sum(marine area), sum(protected area), ratio
am_spp_mpas <- am_spp_cells %>%
  group_by(speciesid) %>%
  summarize(cellarea_tot = sum(cellarea), 
            ocean_range  = sum(oceanarea), 
            marine_range = sum(marine_area),
            mpa1990 = sum(y1990_area),
            mpa1995 = sum(y1995_area),
            mpa2000 = sum(y2000_area),
            mpa2005 = sum(y2005_area),
            mpa2010 = sum(y2010_area),
            mpa2014 = sum(y2014_area),
            mpa2015 = sum(y2015_area),
            pct1990 = mpa1990/marine_range,
            pct1995 = mpa1995/marine_range,
            pct2000 = mpa2000/marine_range,
            pct2005 = mpa2005/marine_range,
            pct2010 = mpa2010/marine_range,
            pct2014 = mpa2014/marine_range,
            pct2015 = mpa2015/marine_range)
```

area_mpa <- cell_df %>%
  summarize(marine_area  = sum(marine_area), 
            mpa_tot_1990 = sum(y1990_area), 
            mpa_tot_1995 = sum(y1995_area), 
            mpa_tot_2000 = sum(y2000_area), 
            mpa_tot_2005 = sum(y2005_area), 
            mpa_tot_2010 = sum(y2010_area), 
            mpa_tot_2014 = sum(y2014_area), 
            mpa_tot_2015 = sum(y2015_area))
#   marine_area  mpa_tot_1990  mpa_tot_1995  mpa_tot_2000  mpa_tot_2005   mpa_tot_2010   mpa_tot_2014
#   362,657,307       351,592       532,098       748,172       932,817      1,137,142      2,885,465

mpa_klein_am2014_filt <- mpa_klein_am2014 %>%
  filter(probability >= thresh)

#   cat2014 n_tot n_cat pct_cat
#       gap 17348   378     2.2
#      0-2% 17348  8859    51.2
#      2-5% 17348  6728    38.7
#     5-10% 17348  1096     6.3
#      >10% 17348   287     1.7

### Redo analysis using 50% threshold, but new AquaMaps database
### * select just 2014 and 2015, at 50% threshold, and group by
###   same categories that the original paper did.
mpa_klein_am2015 <- am_spp_mpas %>%
  select(speciesid, marine_range, mpa2014, mpa2015, pct2014, pct2015)

@@

#   cat2014 n_tot n_cat pct_cat
#       gap 22889   552    2.4
#      0-2% 22889 11546   50.4
#      2-5% 22889  8947   39.1
#     5-10% 22889  1456    6.4
#      >10% 22889   388    1.7

#   cat2015 n_tot n_cat pct_cat
#       gap 22889   552     2.4
#      0-2% 22889 11520    50.3
#      2-5% 22889  8960    39.1
#     5-10% 22889  1466     6.4
#      >10% 22889   391     1.7

@@

| study           | Protected areas    | Area (km^2^) | Gap  |  0–2% |  2–5% | 5–10% | >10% |
| :-------------: | :----------------- | :----------: | :--: | :---: | :---: | :---: | :--: |
| Klein 2015      | IUCN I-IV, Sp. Mar |   2,967,898  | 1.4% | 42.3% | 46.8% |  6.9% | 2.6% |
| am2014, mpa2014 | IUCN I-IV, Sp. Mar |   2,885,465  | 2.2% | 51.1% | 38.8% |  6.3% | 1.7% |
| am2015, mpa2015 | IUCN I-IV, Sp. Mar |   2,889,758  | 2.4% | 50.3% | 39.1% |  6.4% | 1.7% |
| am2015, mpa2014 | IUCN I-IV, Sp. Mar |   2,885,465  | 2.4% | 50.4% | 39.1% |  6.4% | 1.7% |
| am2015, mpa2000 | IUCN I-IV, Sp. Mar |     748,172  | 3.5% | 91.6% |  4.4% |  0.4% | 0.1% |
| am2015, mpa1990 | IUCN I-IV, Sp. Mar |     351,592  | 4.0% | 94.2% |  1.6% |  0.1% | 0.0% |

``` {r check results for 1990 and 2000}
mpa_spp <- am_spp_mpas %>%
  select(speciesid, marine_range, mpa1990, mpa2000, pct1990, pct2000)

@@

@@

#   cat1990 n_tot n_cat pct_cat
#       gap 22889   925     4.0
#      0-2% 22889 21569    94.2
#      2-5% 22889   376     1.6
#     5-10% 22889    18     0.1
#      >10% 22889     1     0.0
# 
#   cat2000 n_tot n_cat pct_cat
#       gap 22889   801     3.5
#      0-2% 22889 20973    91.6
#      2-5% 22889  1004     4.4
#     5-10% 22889    92     0.4
#      >10% 22889    19     0.1

@@

>>>>

@@

dir_rast <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters')
rast_mpa_mar_file <- file.path(dir_rast, 'rast_mpa_mar_i-iv.tif')
# rast_mpa_mar_file <- file.path(dir_rast, 'wdpa_all_mar_klein_i-iv.tif')
rast_mar_file <- file.path(dir_rast, 'rast_marine.tif')

if(!file.exists(rast_mpa_mar_file)) {
#   rast_mpa_file <- file.path(dir_rast, 'WDPA_poly_Jan2015_mpa_i-iv-des.tif')
   rast_mpa_file <- file.path(dir_rast, 'WDPA_poly_Jan2015_mpa_i-iv-all.tif')

@@

rast_loiczid_01_file <- file.path(dir_rast, 'rast_loiczid_01.tif')
if(!file.exists(rast_loiczid_01_file)) {
  rast_loiczid_01 <- resample(rast_loiczid, rast_mpa_mar, 
                            method = 'ngb',
                            progress = 'text',
                            filename = rast_loiczid_01_file)
} else {
  rast_loiczid_01 <- raster(rast_loiczid_01_file)
}
```

@@

### Count marine cells in each LOICZID cell
mar_df_file <- file.path(dir_git, 'mpa_analysis/celldata/loiczid_marine.csv')
if(!file.exists(mar_df_file)) {
  is_marine_df <- zonal(rast_mar, 
                        z = rast_loiczid_01, 
                        fun = 'count',
                        progress = 'text') %>%
    as.data.frame() %>%
    rename(loiczid = zone)
  write_csv(is_marine_df, mar_df_file)
}

### Count all protected cells in each LOICZID cell
### NOTE: loses status_yr values - so filter by year for each iteration
yr_list <- c('y1990', 'y1995', 'y2000', 'y2005', 'y2010', 'y2014', 'y2015')
mpa_cvs_files <- c(file.path(dir_git, 
                             sprintf('mpa_analysis/celldata/loiczid_mpa_i-iv_%s.csv', 
                                     yr_list)))
# mpa_cvs_files <- c(file.path(dir_git, 
#                              sprintf('mpa_analysis/celldata/loiczid_mpa_klein_i-iv_%s.csv', 
#                                      yr_list)))
yr_list1 <- yr_list[!file.exists(mpa_cvs_files)]

if(length(yr_list1) > 0) {
  message('Creating marine protected area rasters for years:')
  message('  ', paste(yr_list1, collapse = ', '))
  for (yr in yr_list1) { # yr <- yr_list1[1]
    message('filtering MPAs to those designated prior to ', str_replace(yr, 'y', ''))
    rast_is_mpa <- rast_mpa_mar
    data_year <- as.integer(str_replace(yr, 'y', ''))
    values(rast_is_mpa)[values(rast_mpa_mar) >= data_year] <- NA
      ### NOTE: 2015 MPA dataset includes up to 2014; 2014 dataset goes
      ###   through 2013; etc.
    message('counting protected cells per LOICZID for ', yr)
    is_mpa_df <- zonal(rast_is_mpa, 
                       z = rast_loiczid_01, 
                       fun = 'count',
                       progress = 'text') %>%
      as.data.frame() %>%
      rename(loiczid = zone)
    write_csv(is_mpa_df, 
              mpa_cvs_files[yr_list == yr])
  }
}
### These files get loaded later; no need to hog memory by loading them here.

@@

dir_am_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/v2015/csv/')
yr_list <- c('y1990', 'y1995', 'y2000', 'y2005', 'y2010', 'y2014', 'y2015')

### get AquaMaps file of cells to translate cell IDs to area and ocean area
hcaf <- read_csv(file.path(dir_am_data, 'hcaf_truncated.csv')) %>%
  select(loiczid, cellarea, oceanarea)

### attach the AquaMaps areas to the calculated percentages of marine area
### note: AquaMaps calculated areas differ from my calculated areas - see below
cell_df <- read_csv(file.path(dir_git, 'mpa_analysis/celldata/loiczid_marine.csv'),
  col_types = 'dd') %>%
  rename(marine_ct = count) %>%
  left_join(hcaf, by = 'loiczid') %>%
  mutate(marine_pct  = marine_ct/2500,
         marine_area = marine_pct * cellarea) %>%
  select(-marine_ct)

### load each analysis year (in yr_list) and join them all together
### to note protection status year of each cell.

### Use this to point toward the 'i-iv-all' files (for all STATUS) or
### 'i-iv-des' (for STATUS == 'Designated')
mpa_file_base <- 'mpa_analysis/celldata/loiczid_mpa_i-iv_%s.csv'
# mpa_file_base <- 'mpa_analysis/celldata/loiczid_mpa_klein_i-iv_%s.csv'
# mpa_file_base <- 'mpa_analysis/celldata/loiczid_mpa_i-iv-des_%s.csv'

for (yr in yr_list) { # yr <- yr_list[1]
  fn <- file.path(dir_git, sprintf(mpa_file_base, yr))
  df <- read_csv(fn, col_types = 'dd')
  names(df)[2] <- yr
  cell_df <- cell_df %>%
    left_join(df, by = 'loiczid')
  cell_df[ , paste(yr, '_pct', sep = '')]  <- cell_df[ , yr]/2500
  cell_df[ , paste(yr, '_area', sep = '')] <- cell_df[ , paste(yr, '_pct', sep = '')] * cell_df$cellarea

@@

# recreate Klein results using aquamaps 2014 data

area_mpa <- cell_df %>%
  summarize(marine_area  = sum(marine_area), 
            mpa_tot_1990 = sum(y1990_area), 
            mpa_tot_1995 = sum(y1995_area), 
            mpa_tot_2000 = sum(y2000_area), 
            mpa_tot_2005 = sum(y2005_area), 
            mpa_tot_2010 = sum(y2010_area), 
            mpa_tot_2014 = sum(y2014_area), 
            mpa_tot_2015 = sum(y2015_area))

mpa_klein_am2014 <- mpa_klein_am2014 %>%
  filter(probability >= thresh)

### Not filtering on status, but all in shapefile are status == 'Designated'
#   cat2014 n_tot n_cat pct_cat
#       gap 17344   282     1.6
#      0-2% 17344  7728    44.6
#      2-5% 17344  7738    44.6
#     5-10% 17344  1188     6.8
#      >10% 17344   408     2.4

# run analysis using 2015 aquamaps

@@

### group by speciesid, then summarize sum(marine area), 
### sum(protected area), ratio
am_spp_mpas_filt <- am_spp_cells_filt %>%
  group_by(speciesid) %>%
  summarize(cellarea_tot = sum(cellarea), 
            ocean_range  = sum(oceanarea), 
            marine_range = sum(marine_area),
            mpa1990 = sum(y1990_area),
            mpa1995 = sum(y1995_area),
            mpa2000 = sum(y2000_area),
            mpa2005 = sum(y2005_area),
            mpa2010 = sum(y2010_area),
            mpa2014 = sum(y2014_area),
            mpa2015 = sum(y2015_area),
            pct1990 = mpa1990/marine_range,
            pct1995 = mpa1995/marine_range,
            pct2000 = mpa2000/marine_range,
            pct2005 = mpa2005/marine_range,
            pct2010 = mpa2010/marine_range,
            pct2014 = mpa2014/marine_range,
            pct2015 = mpa2015/marine_range)
```

``` {r redo klein for new AquaMaps with 2015 mpas}

area_mpa <- cell_df %>%
  summarize(marine_area  = sum(marine_area), 
            mpa_tot_1990 = sum(y1990_area), 
            mpa_tot_1995 = sum(y1995_area), 
            mpa_tot_2000 = sum(y2000_area), 
            mpa_tot_2005 = sum(y2005_area), 
            mpa_tot_2010 = sum(y2010_area), 
            mpa_tot_2014 = sum(y2014_area), 
            mpa_tot_2015 = sum(y2015_area))

#   marine_area mpa_tot_1990 mpa_tot_1995 mpa_tot_2000 mpa_tot_2005 mpa_tot_2010 mpa_tot_2014 mpa_tot_2015
#         (dbl)        (dbl)        (dbl)        (dbl)        (dbl)        (dbl)        (dbl)        (dbl)
# 1   362657307     353355.3     534137.2     750217.9     935137.1      1140272      2888600      2896875

### Redo analysis using 50% threshold, but new AquaMaps database
### * select just 2014 and 2015, at 50% threshold, and group by
###   same categories that the original paper did.
mpa_klein_am2015 <- am_spp_mpas_filt %>%
  select(speciesid, marine_range, mpa2014, mpa2015, pct2014, pct2015)

@@

# > mpa_klein_sum2014
#   cat2014 n_tot n_cat pct_cat
#       gap 22885   797     3.5
#      0-2% 22885 11038    48.2
#      2-5% 22885  8713    38.1
#     5-10% 22885  1650     7.2
#      >10% 22885   687     3.0
# 
# > mpa_klein_sum2015
#   cat2015 n_tot n_cat pct_cat
#       gap 22885   797     3.5
#      0-2% 22885 10990    48.0
#      2-5% 22885  8747    38.2
#     5-10% 22885  1661     7.3
#      >10% 22885   690     3.0

@@

``` {r check results for 1990 through 2010 at 50%}
mpa_spp <- am_spp_mpas_filt %>%
  select(speciesid, marine_range, mpa1990:mpa2010, pct1990:pct2010)

@@

mpa_spp_1995 <- mpa_spp %>%
  group_by(cat1995) %>%
  summarize(n_tot = nrow(mpa_spp), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))

@@

mpa_spp_2005 <- mpa_spp %>%
  group_by(cat2005) %>%
  summarize(n_tot = nrow(mpa_spp), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))

mpa_spp_2010 <- mpa_spp %>%
  group_by(cat2010) %>%
  summarize(n_tot = nrow(mpa_spp), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))

#   cat1990 n_tot n_cat pct_cat
#       gap 22889   925     4.0
#      0-2% 22889 21569    94.2
#      2-5% 22889   376     1.6
#     5-10% 22889    18     0.1
#      >10% 22889     1     0.0
# 
#   cat1995 n_tot n_cat pct_cat
#       gap 22889   878     3.8
#      0-2% 22889 21111    92.2
#      2-5% 22889   817     3.6
#     5-10% 22889    70     0.3
#      >10% 22889    13     0.1
#
#   cat2000 n_tot n_cat pct_cat
#       gap 22889   801     3.5
#      0-2% 22889 20973    91.6
#      2-5% 22889  1004     4.4
#     5-10% 22889    92     0.4
#      >10% 22889    19     0.1
#
#   cat2005 n_tot n_cat pct_cat
#       gap 22889   608     2.7
#      0-2% 22889 19828    86.6
#      2-5% 22889  2174     9.5
#     5-10% 22889   205     0.9
#      >10% 22889    74     0.3
# 
#   cat2010 n_tot n_cat pct_cat
#       gap 22889   578     2.5
#      0-2% 22889 18670    81.6
#      2-5% 22889  3278    14.3
#     5-10% 22889   288     1.3
#      >10% 22889    75     0.3

argh <- am_spp_mpas_filt %>%
  select(-(mpa1990:mpa2015)) %>%
  gather(key = year, value = pct, pct1990:pct2015) %>%
  mutate(year = str_replace(year, 'pct', '')) %>%
  group_by(year)

hist_plot <- ggplot(data = argh, aes(x = pct, color = year)) + 
  geom_density() +
  scale_x_continuous(limits = c(0, .05))
#  scale_x_log10()

hist_plot
ggsave(file.path(dir_git, 'figures/mpa_analysis_dist.png'))
```

| study           | Protected areas    | Area (km^2^) | Gap  |  0–2% |  2–5% | 5–10% | >10% |
| :-------------: | :----------------- | :----------: | :--: | :---: | :---: | :---: | :--: |
| Klein et al 2015| IUCN I-IV, Sp. Mar |   2,967,898  | 1.4% | 42.3% | 46.8% |  6.9% | 2.6% |
| am2014, mpa2014 | IUCN I-IV, Sp. Mar |   2,915,745  | 1.6% | 44.6% | 44.6% |  6.8% | 2.4% |
| .               | .                  | .            | .    | .     | .     | .     | .    |
| am2015, mpa2015 | IUCN I-IV, Sp. Mar |   2,889,758  | 2.4% | 50.3% | 39.1% |  6.4% | 1.7% |
| am2015, mpa2014 | IUCN I-IV, Sp. Mar |   2,885,465  | 2.4% | 50.4% | 39.1% |  6.4% | 1.7% |
| am2015, mpa2010 | IUCN I-IV, Sp. Mar |   1,137,142  | 2.5% | 81.6% | 14.3% |  1.3% | 0.3% |
| am2015, mpa2005 | IUCN I-IV, Sp. Mar |     932,817  | 2.7% | 86.6% |  9.5% |  0.9% | 0.3% |
| am2015, mpa2000 | IUCN I-IV, Sp. Mar |     748,172  | 3.5% | 91.6% |  4.4% |  0.4% | 0.1% |
| am2015, mpa1995 | IUCN I-IV, Sp. Mar |     532,098  | 3.8% | 92.2% |  3.6% |  0.3% | 0.1% |
| am2015, mpa1990 | IUCN I-IV, Sp. Mar |     351,592  | 4.0% | 94.2% |  1.6% |  0.1% | 0.0% |

max protected species = Fis-160882

``` {r iucn_only}
### compare results using only IUCN species maps
dir_iucn_data <- '/var/data/ohi/git-annex/globalprep/_raw_data/iucn_spp'
iucn_cells_file <- file.path(dir_iucn_data, 'iucn_cells_2015.csv')
iucn_spp_cells <- read_csv(iucn_cells_file, col_types = 'cddddc') %>%
  as.data.frame(stringsAsFactors = FALSE)

# iucn_spp_cells <- read_csv(file.path(dir_am_data, 'hcaf_sp_native_trunc.csv'))
iucn_spp_cells <- iucn_spp_cells %>%
  left_join(cell_df, by = 'loiczid')

### group by speciesid, then summarize sum(marine area), sum(protected area), ratio
iucn_spp_sum <- iucn_spp_cells %>%
  group_by(iucn_sid) %>%
  summarize(cellarea_tot = sum(cellarea), 
            ocean_range  = sum(oceanarea), 
            marine_range = sum(marine_area),
            mpa2015 = sum(y2015_area),
            pct2015 = mpa2015/marine_range)

iucn_spp_sum <- iucn_spp_sum %>%
  mutate(cat2015 = ifelse(pct2015 == 0,                       'gap', NA),
         cat2015 = ifelse(pct2015 > 0    & pct2015 <= 0.02,  '0-2%', cat2015),
         cat2015 = ifelse(pct2015 > 0.02 & pct2015 <= 0.05,  '2-5%', cat2015),
         cat2015 = ifelse(pct2015 > 0.05 & pct2015 <= 0.10, '5-10%', cat2015),
         cat2015 = ifelse(pct2015 > 0.10,                    '>10%', cat2015))

iucn_spp_sum2 <- iucn_spp_sum %>%
  group_by(cat2015) %>%
  summarize(n_tot = nrow(iucn_spp_sum), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))

@@

``` {r combine IUCN and AM with 0% thresh, AM on top}

### use iucn_spp_sum from above, and create am_spp_sum for 0% thresh.
iucn_spp_sum

am_spp_mpas_0 <- am_spp_cells %>%
  group_by(speciesid) %>%
  summarize(cellarea_tot = sum(cellarea), 
            ocean_range  = sum(oceanarea), 
            marine_range = sum(marine_area),
            mpa2015 = sum(y2015_area),
            pct2015 = mpa2015/marine_range)

am_spp_sum <- am_spp_sum %>%
  mutate(cat2015 = ifelse(pct2015 == 0,                       'gap', NA),
         cat2015 = ifelse(pct2015 > 0    & pct2015 <= 0.02,  '0-2%', cat2015),
         cat2015 = ifelse(pct2015 > 0.02 & pct2015 <= 0.05,  '2-5%', cat2015),
         cat2015 = ifelse(pct2015 > 0.05 & pct2015 <= 0.10, '5-10%', cat2015),
         cat2015 = ifelse(pct2015 > 0.10,                    '>10%', cat2015))

# combine by bringing in species list with IUCN SID mapped to AM SID;
# use all with AM SID, and then add in IUCN spp that are not mapped to an AM SID

``` {r get_iucn_cells}
reload = TRUE
  dir_iucn_data <- '/var/data/ohi/git-annex/globalprep/_raw_data/iucn_spp'
  iucn_cells_file <- file.path(dir_iucn_data, 'iucn_cells_2015.csv')
  if(!file.exists(iucn_cells_file) | reload) {
    message(sprintf('Building IUCN species to cell table.  This might take a few minutes.\n'))
    iucn_map_files      <- file.path(dir_anx, 'iucn_intersections', list.files(file.path(dir_anx, 'iucn_intersections')))
    iucn_cells_spp_list <- lapply(iucn_map_files, read.csv) # read each into dataframe within a list
    iucn_cells_spp      <- bind_rows(iucn_cells_spp_list)   # combine list of dataframes to single dataframe
    # This creates a full data frame of all IUCN species, across all species groups, for all cells.
    # Probably big...
    names(iucn_cells_spp) <- tolower(names(iucn_cells_spp))
    write_csv(iucn_cells_spp, iucn_cells_file)
  } else {
    iucn_cells_spp <- read_csv(iucn_cells_file, col_types = 'cddddc') %>%
      as.data.frame(stringsAsFactors = FALSE)
  }

}
```

### 2016-03-01 mpa_analysis/wdpa_prep.Rmd
meh
let's get some of this up on line...

### import WDPA database polygons from saved .shp
dir_wdpa   <- file.path(dir_N, 'git-annex/globalprep/_raw_data/wdpa_mpa/WDPA_Jan2015_Public')
lyr_wdpa    <- 'WDPA_poly_Jan2015'
lyr_mpa     <- 'WDPA_poly_Jan2015_mpa'
lyr_mpa_mar <- 'WDPA_poly_Jan2015_mpa_mar'

@@

  poly_mpa <- poly_wdpa[poly_wdpa@data$STATUS == 'Designated', ]
  poly_mpa <- poly_mpa[poly_mpa@data$prot_code <= 4, ]

  poly_mpa_marine <- poly_mpa[poly_mpa@data$MARINE == 1, ]

  writeOGR(poly_mpa, dsn = dir_wdpa, layer = lyr_mpa,
           driver = 'ESRI Shapefile', overwrite_layer = TRUE)
  writeOGR(poly_mpa_marine, dsn = dir_wdpa, layer = lyr_mpa_mar,
           driver = 'ESRI Shapefile', overwrite_layer = TRUE)
}

poly_mpa <- readOGR(dir_wdpa, lyr_mpa, stringsAsFactors = FALSE)
# poly_mpa <- readShapePoly(fn = file.path(dir_wdpa, lyr_mpa),
#                           proj4string = crs_wgs84,
#                           stringsAsFactors = FALSE)

### create simplified polygon
polys_mpa_area <- lapply(poly_mpa@polygons, function(x) x@area)
poly_mpa1 <- poly_mpa[polys_mpa_area > 1e-5, ]
poly_mpa1 <- poly_simplify(poly_mpa1, thresh = 1e-5, verbose = TRUE)

### sort by STATUS_YR, descending, so gdal_rasterize will assign newer
### values then overwrite (where applicable) with older values
poly_mpa1 <- poly_mpa1[order(poly_mpa1@data$STATUS_YR, decreasing = TRUE), ]

# writeOGR(poly_mpa1, dsn = dir_wdpa,
#          layer = paste(lyr_mpa, '1', sep = ''),
#          driver = 'ESRI Shapefile')
maptools::writePolyShape(poly_mpa1, 
                         fn = file.path(dir_wdpa, paste(lyr_mpa, '1', sep = '')))
### copy .prj:
file.copy(from = file.path(dir_wdpa, paste(lyr_mpa, '.prj', sep = '')), 
          to = file.path(dir_wdpa, paste(lyr_mpa, '1.prj', sep = '')))

@@

### Get source from WDPA_MPA data location
src <- path.expand(file.path(dir_wdpa, paste(lyr_mpa, '1.shp', sep = '')))
### gdal_rasterize to temp directory
tmp_dst <- file.path(dir_rast, 'tmp', paste(lyr_mpa, '1.tif', sep = '')) %>%
  path.expand
### then writeRaster to final directory, to save smaller file
fnl_dst <- file.path(dir_rast, paste(lyr_mpa, '1.tif', sep = '')) %>%
  path.expand

@@

>>>>

### import WDPA database polygons from saved .shp
dir_wdpa   <- file.path(dir_N, 'git-annex/globalprep/_raw_data/wdpa_mpa/WDPA_Jan2015_Public')
lyr_wdpa    <- 'WDPA_poly_Jan2015'
lyr_mpa     <- 'WDPA_poly_Jan2015_mpa_i-iv-all'

@@

  poly_mpa <- poly_wdpa[poly_wdpa@data$prot_code <= 4, ]
  #  poly_mpa <- poly_mpa[poly_mpa@data$STATUS == 'Designated', ]

  ### sort by STATUS_YR, descending, so gdal_rasterize will assign newer
  ### values then overwrite (where applicable) with older values
  poly_mpa <- poly_mpa[order(poly_mpa@data$STATUS_YR, decreasing = TRUE), ]

  maptools::writePolyShape(poly_mpa, 
                           fn = file.path(dir_wdpa, lyr_mpa))
  ### copy .prj:
  file.copy(from = file.path(dir_wdpa, paste(lyr_wdpa, '.prj', sep = '')), 
            to = file.path(dir_wdpa, paste(lyr_mpa, '.prj', sep = '')))

# writeOGR(poly_mpa, dsn = dir_wdpa, layer = lyr_mpa,
#            driver = 'ESRI Shapefile', overwrite_layer = TRUE)
#   writeOGR(poly_mpa_marine, dsn = dir_wdpa, layer = lyr_mpa_mar,
#            driver = 'ESRI Shapefile', overwrite_layer = TRUE)
}

@@

### Get source from WDPA_MPA data location
src <- path.expand(file.path(dir_wdpa, paste(lyr_mpa, '.shp', sep = '')))
### gdal_rasterize to temp directory
tmp_dst <- file.path(dir_rast, 'tmp', paste(lyr_mpa, '.tif', sep = '')) %>%
  path.expand
### then writeRaster to final directory, to save smaller file
fnl_dst <- file.path(dir_rast, paste(lyr_mpa, '.tif', sep = '')) %>%
  path.expand

@@

``` {r set_up_wdpa_mpa_polys_klein}

### import WDPA database polygons from saved .shp
dir_rast <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters')
dir_wdpa   <- file.path(dir_rast, 'WDPA2014_mar_klein')
lyr_wdpa    <- 'wdpa_all_mar_klein'
lyr_mpa     <- 'wdpa_all_mar_klein_i-iv'

  prot_code_df <- data.frame('IUCN_CAT'  = c('Ia', 'Ib', 'II', 'III', 'IV', 'V', 'VI'),
                             'prot_code' = c(  1,    1,    2,     3,    4,   5,    6))

  poly_wdpa@data <- poly_wdpa@data %>%
    left_join(prot_code_df, by = 'IUCN_CAT') %>%
    mutate(prot_code = ifelse(is.na(prot_code), 7, prot_code))

  poly_mpa <- poly_wdpa[poly_wdpa@data$prot_code <= 4, ]

  # poly_mpa <- poly_mpa[poly_mpa@data$STATUS == 'Designated', ]

  if(poly_mpa@proj4string@projargs != p4s_wgs84) {
    message('Polygons not in WGS84 projection.  \n  Polygon projection: ', 
            poly_mpa@proj4string@projargs)
    poly_mpa <- spTransform(poly_mpa, crs_wgs84)
  }

  poly_mpa <- poly_mpa[order(poly_mpa1@data$STATUS_YR, decreasing = TRUE), ]

  writeOGR(poly_mpa, dsn = dir_wdpa, layer = lyr_mpa,
             driver = 'ESRI Shapefile', overwrite_layer = TRUE)
  #   writeOGR(poly_mpa_marine, dsn = dir_wdpa, layer = lyr_mpa_mar,
  #            driver = 'ESRI Shapefile', overwrite_layer = TRUE)
}
